If the Null Fits, You Must Omit: Ubiquitous False Detections of Group Polarization

### 2025-08-06 GroupPolarizationStatmod/paper/intro.qmd
an update to Intro.qmd from June; merge with the rest by compiling intro.tex and including with the rest of the latex draft

- Having established a model-based definition of null-consistent detection, we now apply it to published group polarization experiments.
- By combining this logic with simulations seeded from published data, we estimate false discovery rates across conditions.
- This isn’t a judgment about whether past detections were true. It’s a demonstration of how to estimate false detection rates properly—using models that actually fit the structure of the data. We calculate FWER and FDR for the original study designs at three effect size levels—small, medium, and large—based on Cohen’s guidelines [@CohenBook1988, Ch. 2.2.3] and Bayesian methods from Liddell and Kruschke [@Liddell2018]. The verdict is clear: even if these studies had used better models, most still wouldn’t have had the power to rule out false positives. That makes replication a weak signal of reliability. If future group polarization studies want to produce meaningful results, they need stronger designs—and that starts with a clean break from outdated statistical habits.

>>>>

- Having established a model-based definition of null-consistent detection, we now apply it to published group polarization experiments.
- By combining this logic with simulations seeded from published data, we estimate false discovery rates across conditions.
- This isn’t a judgment about whether past detections were true. It’s a demonstration of how to estimate false detection rates properly—using models that actually fit the structure of the data. We calculate FWER and FDR for the original study designs at three effect size levels—small, medium, and large—based on Cohen’s guidelines [@CohenBook1988, Ch. 2.2.3] and Bayesian methods from Liddell and Kruschke [@Liddell2018]. The verdict is clear: even if these studies had used better models, most still wouldn’t have had the power to rule out false positives. That makes replication a weak signal of reliability. If future group polarization studies want to produce meaningful result need stronger designs—and that starts with a clean break from outdated statistical habits.
